---
title: "Las Vegas Restaurants"
author: "Claire Wang"
date: 2018-05-20
output: 
  github_document:
    toc: true
    toc_depth: 6
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = '100%')
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)

# Files
file_yelp <-

```

Every year, Yelp releases a public dataset of local businesses for students to explore and use in innovative research. The data includes information on businesses, ratings, and reviews. More information on Yelp's data and student challenge can be found [here](https://www.yelp.com/dataset/challenge). Since the data changes each year, please download the 2017 dataset on [Box](https://stanford.box.com/s/eijybzf2mhbfrahlxlfauenxb4rgiwu0). 

For this challenge you will explore businesses in Las Vegas, focusing on restaurants and food-related businesses (e.g. bakeries, grocery stores, etc.). You will do some preliminary EDA, then finish by creating restaurant maps for Las Vegas (by cuisine type) and ultimately automating this map-making process. 


## Detailed data

### Download raw data

Download the raw data from [Box](https://stanford.box.com/s/eijybzf2mhbfrahlxlfauenxb4rgiwu0). Read in the csv file as a data frame, then use `glimpse` or `summary` to preview the data. 

```{r}
# solution-begin

df <-
  file_yelp %>%
  read_csv()

# solution-end
```

### Clean data for analysis

__q1.0__ Is this data exclusively for businesses in the U.S., or does it include other countries? Create a dataframe `q1` that only contains businesses in the U.S. (Hint: use variables `state` and `postal_code`.)

```{r}
# solution-begin

q1 <-
  df %>%
  filter(
    !state %in% c("6", "30", "3", "01"),
    str_length(state) == 2,
    str_length(postal_code) == 5
  ) 
# solution-end
```

<!-- solution-begin -->
Exploring the data, we saw that there are states that are numerically coded, so we removed states with the numbers 6, 30, 3, and 01 as values. We also filtered out any states with more or less than 2 letters in the code. We also removed postal codes with more or less than 5 digits. 
<!-- solution-end -->

__q1.1__ Are the states well-represented in this dataset, or is the data concentrated in certain states? To answer this, print out the top 5 states with the most businesses in this dataset. 

```{r}
# solution-begin
q1 %>%
  count(state) %>%
  select(state, n_biz = n) %>%
  arrange(desc(n_biz)) %>%
  top_n(n = 5)
# solution-end

```

<!-- solution-begin -->
Arizona has the most observations, with more than 52,000 businesses. Nevada is second, with just under 33,000. North Carolina, Ohio, and Pennsylvania are the next three most represented states.   
<!-- solution-end -->


__q1.2__ From Q1.1, we saw that Arizona and Nevada have the most businesses. We will focus on businesses in Nevada - specifically, businesses in Las Vegas. Create a new dataset, `q1.2`, that only includes businesses in Las Vegas. 

```{r}
# solution-begin
q1.2 <-
  q1 %>%
  filter(city == "Las Vegas")
# solution-end
```

## Create restaurant data

__q2.0__ We are interested in creating a restaurant map, so we want to filter out our data further to only include food-related businesses. 

From `q1.2`, create a new dataframe `q2` that only contains businesses whose category includes the string "Restaurants" or "Food".

```{r}
# solution-begin
q2 <-
  q1.2 %>%
  filter(
    str_detect(categories, "Restaurants|Food")
  )
# solution-end
```

__q2.1__ You may have noticed that each restaurant's `categories` include multiple values. For example, a restaurant's category might be "Restaurants, American, Burgers, Fast Food". We want to have only one label per restaurant. 

This process is a bit subjective, but for the purposes of this challenge we will do the following. Using `q2`, create a dataframe `q2.1` using the following steps:

* Use the function `separate_rows()` to create a new observation for each label in `categories` for each restaurant. You may want to define the `sep = ` parameter. 
* Since we have multiple observations now for the same business, create a new variable called `row_num` that sequentially numbers the rows for each individual business. `row_num` should start over at 1 for each new business. We will use this variable later to ensure we end up with one observation per restaurant. 
* All food-related businesses have generic values in their `categories`, such as "Food" or "Restaurants". We do __not__ want these generic labels as our ultimate category; instead, we are interested in _specific_ `categories` labels, such as "Mexican" or "Bakeries". To solve this, remove any observations whose `categories` value is in the `generic_food_labels` vector defined below.
* Now, to get one observation per business, keep only the minimum row_num for each business (`row_num == min(row_num)`). 

You can do all of this with one pipe. Don't create multiple intermediate datasets.

```{r}
generic_food_labels <- 
  c(
    "Restaurants", 
    "Food", 
    "Food Trucks", 
    "Food Delivery Services",
    "Food Stands", 
    "Event Planning & Services", 
    "Food Court", 
    "Caterers",
    "Shopping",
    "Arts & Entertainment",
    "Car Wash",
    "Gas Stations",
    "Automotive",
    "Local Flavor",
    "Bookstores",
    "Nightlife",
    "Venues & Event Spaces",
    "Business Consulting",
    "Hotels & Travel",
    "Social Clubs",
    "Party & Event Planning",
    "Food Tours"
  )

# solution-begin
q2.1 <-
  q2 %>%
  separate_rows(categories, sep = ";") %>%
  group_by(business_id) %>%
  mutate(row_num = row_number()) %>%
  filter(!categories %in% generic_food_labels) %>%
  filter(row_num == min(row_num)) %>%
  ungroup()
# solution-end
```

## Restaurant EDA

Now we have a clean dataset of Las Vegas restaurants (q2.1) that we can explore.

__q3.0__ Is a restaurant's rating related to its number of reviews? Create a visualization that answers this question and interpret.

```{r}
# solution-begin
q2.1 %>%
  ggplot(mapping = aes(x = stars, y = review_count)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(
    title = "Restaurant Ratings by Review Count",
    subtitle = "Restaurants have higher review counts as their rating goes up, up to 4 stars",
    x = "Restaurant Average Rating",
    y = "Number of Reviews"
  )

# solution-end
```

<!-- solution-begin -->
Restaurants with higher average ratings have higher review counts until 4 stars. After 4 stars, the number of reviews decreases as rating increases. It could be that most users rate 3.5-4 stars and are less hesitant to give a restaurant a 1, 2, or 5 star rating.   
<!-- solution-end -->


__q3.1__ Create a vector called `top_15_categ` that includes the top 15 most common restaurant types in Las Vegas. What are the top 5 most common types of restaurants in Las Vegas?

```{r}
# solution-begin
top_15_categ <-
  q2.1 %>%
  count(categories, sort = TRUE) %>%
  top_n(n = 15) %>%
  pull(categories)

top_15_categ[1:5]
# solution-end

```

__q3.2__ Using dataframe q2.1 and top_15_categ, create a visualization that explores the rating distributions for each of the top 15 categories. You can either recreate [this chart](https://imgur.com/a/ttpcdNh) or create a better one of your own!

```{r}
# solution-begin
q2.1 %>%
  filter(categories %in% top_15_categ) %>%
  count(stars, categories) %>%
  ggplot(mapping = aes(x = stars, y = n, color = categories)) +
  geom_line() +
  labs(
    title = "Ratings by Restaurant Type",
    x = "Average Rating",
    y = "Number of Restaurants",
    color = "Restaurant Type",
    caption = "Source: Yelp"
  ) 

# solution-end

```



__q3.3__ Create a visualization of your own to uncover a new insight about restaurants in Las Vegas.
```{r}

```


## Creating a Restaurant Map

Our Yelp data uses latitude and longitude to identify a business's location. We will use two R libraries, `ggmap` and `mapproj`, to create our maps, as they use latitude and longitude data. You may want to move the library calls below to the top of the R markdown file for style purposes. 

```{r}
library(ggmap)
library(mapproj)
```


__q4.0__ `ggmap` has pre-created maps that we can load and use. Create a variable `lv_map` that is a map of Las Vegas (Hint: use `get_map()`, specifying `location` as Las Vegas and `zoom` to get a good view). Use `ggmap(lv_map)` to view the map you have created. 

```{r}
# solution-begin

lv_map <- 
  get_map(
    location = 'Las Vegas',
    zoom = 12
  )
# solution-end

ggmap(lv_map)


```



__q4.1__ We can plot our restaurant data directly on the map of Las Vegas by using `ggmap(lv_map) + geom_point(...)`. Create a dataframe `q4.1` that only consists of Las Vegas "Burgers" restaurants. Plot the points of `q4.1` over the map of Las Vegas, `lv_map`, to recreate [this burger map](https://imgur.com/a/M8ua0cD).

```{r}
# solution-begin

q4.1 <-
  q2.1 %>%
  filter(categories == "Burgers")

map_output <-
  ggmap(lv_map) +
  geom_point(
    data = q4.1,
    aes(
      x = longitude,
      y = latitude,
      color = factor(format(round(stars, 1), nsmall = 1))
    ),
    alpha = 0.7
  ) +
  labs(
    title = "Burgers Food Map",
    x = NULL,
    y = NULL,
    caption = "Source: Yelp",
    color = "Business Rating"
  )

map_output
# solution-end

```


## Generalizing the Restaurant Map
We were able to create a map for Burgers, but suppose we want to create this same restaurant map for each of the 15 most popular restaurant types that we identified in `top_15_categ`. 

__q5.0__ First, create a function called `create_food_map` that inputs a particular cuisine type (string) and returns the map we created above but for this specific cuisine type (`print(image)` may be helpful to return the image). Remember to update the map's subtitle so that it is customized for whatever cuisine type you are entering in (e.g. "Category: Pizza", "Category: Mexican").   

```{r}
# solution-begin

# Create function
create_food_map <- function(x) {
  # Filter dataset based on x cuisine
  df_cuisine <-
    q2.1 %>%
    filter(categories == x)
  
  # Create map
  map_output <-
    ggmap(lv_map) +
    geom_point(
      data = df_cuisine,
      aes(
        x = longitude,
        y = latitude,
        color = factor(format(round(stars, 1), nsmall = 1))
      ),
      alpha = 0.7
    ) +
    labs(
      title = "Las Vegas Restaurant Map",
      subtitle = str_c("Category: ", df_cuisine$categories),
      x = NULL,
      y = NULL,
      caption = "Source: Yelp",
      color = "Business Rating"
    )
  
  print(map_output)
}

# solution-end
```

__q5.1__ Using the function you just created and `top_15_categ`, write one line of code that will create a food map for each of the cuisine types in `top_15_categ`. When you run this code, the output should be a total of 15 restaurant maps, each customized for a specific cuisine type.  

```{r}
# solution-begin
walk(top_15_categ, create_food_map)
# solution-end
```

